\textbf{How open are hybrid journals included in transformative agreements?}

### 2024-09-25 analysis/manuscript.Rmd
Add figure data

```{r cr_stats_overview}
cr_stats <- readr::read_csv(here::here("data", "cr_stats.csv"))
cr_stats_plot <- cr_stats |>
  mutate(cat = case_when(
    is.na(src) & oa_status == "gold" ~ "gold",
    src == "hoad" ~ "hoad",
    .default = "other"
  )) |>
  group_by(cr_year, cat) |>
  summarise(n = sum(n_articles)) |>
  mutate(cat = factor(cat, levels = c("hoad", "other", "gold"))) |>
  mutate(prop = n / sum(n)) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), n, fill = cat)) +
  geom_bar(
    position = position_stack(reverse = TRUE),
    color = "black",
    stat = "identity"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels =  scales::number_format(big.mark = ",")) +
  scale_fill_manual(
    "Journal",
    values = c(
      "hoad" = "#998ec3",
      "other" = "#f7f7f7",
      "gold" = "#f1a340"
    ),
    labels = c(
      "hoad" = "Hybrid (TA)",
      "other" = "Other",
      "gold" = "Full OA"
    )
  ) +
  theme_minimal() +
  labs(x = "Publication Year", y = "Total Articles") +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title = element_text(size = 10)#,
  #  legend.position = "right",
  #  legend.justification = "right"
  ) 
```

@@

```{r method_fig, fig.asp=0.55, fig.cap="Initial data characteristics. (A) Full open access journals included in transformative agreements by evidence source Directory of Open Access Journals (DOAJ), OpenAlex and the Bielefeld GOLD OA list. (B) Number of articles in Crossref by journal types. The purple bars show the overall article volume of hybrid journals in transformative agreements, which were initially included in the study, in comparision with full open access journals according to OpenAlex. The remainder represents articles in subscription-based journals not covered by transformative agreements."}
(p | cr_stats_plot) + plot_annotation(tag_levels = 'A') +  plot_layout(widths = c(0.6, 0.4))
```

@@

oa_jn_box_df <- left_join(jn_hybrid_oa_growth, jn_year_all) |>
  mutate(prop = oa_all / n) |> 
  filter(!cr_year %in%  c("2017", "2023"))
oa_jn_box <- oa_jn_box_df |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  coord_cartesian(ylim = c(0,0.65)) +
  theme_minimal() +
  labs(y = "OA Uptake per Hybrid Journal", x = "Publication Year")
```

@@

layout = "
EEE
ACD
BCD"
wrap_plots(
  E = guide_area(),
  A = oa_ta_plot,
  B = ta_share_plot,
  C = oa_jn_box,
  D = box_ta_prop,
  design = layout,
  widths = c(0.4, 0.3, 0.3),
  heights = c(0.2, 1, .5),
  guides = "collect") &
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
 #   panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A')
```

@@

publisher_summary_df <- inner_join(oa_publisher, all_publisher) |>
  mutate(prop = value / jn_all)

@@

publisher_ta_prop_plot <- publisher_ta_prop |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_bar(stat = "identity",
           fill = "#56B4E9",
            color = "#000000",
           linewidth = 0.15) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    breaks = c(0, 0.25, 0.5, 0.75),
    limits = c(0, .75),
    labels = scales::percent
  ) +
  facet_wrap( ~ publisher, nrow = 1) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
  theme(
    axis.text.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank()
  )
```

@@

design = "E
A
B
C"
wrap_plots(
  E = guide_area(),
  A = publisher_prop,
  B = publisher_ta_prop_plot,
  C = publisher_boxplot,
  design = design,
  guides = "collect",
  heights = c(0, 0.4, 0.3, 0.7)
) &
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
#    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A') 
```

@@

```{r subject_plots}
subject_trend_plot <- top_level_by_year_oa |>
  filter(top_level != "Multidisciplinary") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  select(cr_year, top_level, contains("frac"),-oa_frac) |>
  pivot_longer(
    cols = c(ta_oa_frac, no_ta_fractions),
    names_to = "type",
    values_to = "value"
  ) |>
  mutate(prop = value / articles_frac) |>
  ggplot(aes(
    gsub("^20", "'", as.character(cr_year)),
    prop,
    fill = type,
    group = type
  )) +
    geom_area(color = "#000000",
            linewidth = 0.15) +
  facet_wrap( ~ top_level, nrow = 1) +
  scale_fill_manual(
    NULL,
    values = c(
      no_ta_fractions = "#b3b3b3a0",
      ta_oa_frac = "#56B4E9"
    ),
    labels = c(no_ta_fractions = "Other", ta_oa_frac = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    axis.text.x = element_blank()
  )

subject_ta_trend_plot <- top_level_by_year_oa |>
  filter(top_level != "Multidisciplinary") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  mutate(prop = ta_oa_frac / oa_frac) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_bar(stat = "identity", fill = "#56B4E9", color = "#000000",
            linewidth = 0.15) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
                     labels = scales::percent) +
  facet_wrap(~top_level, nrow = 1) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
  theme(
    axis.text.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank())
subject_box_trend <- subject_df |>
   # with at least one
  select(-ajcs, -subject_area, -esac_publisher) |>
  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
  filter(value > 0) |>
  filter(top_level != "Multidisciplinary", !cr_year %in% c("2017", "2023")) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE, varwidth = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
  coord_cartesian(ylim = c(0,0.55)) +
  facet_wrap(~top_level, nrow = 1) +
  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
  theme_minimal() +
    theme(strip.background = element_blank(),
    strip.text.x = element_blank()
    )
```

@@

design = "E
A
B
C"
wrap_plots(
  E = guide_area(),
  A = subject_trend_plot,
  B = subject_ta_trend_plot,
  C = subject_box_trend,
  design = design,
  guides = "collect",
  heights = c(0, 0.4, 0.3, 0.7)
) &
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A') 
```

@@

top_20_plot_df <- 
  country_multiple_df |> 
  mutate(no_ta = oa_total - ta_oa_total) |>
  rename(ta_oa = ta_oa_total) |>
  pivot_longer(
  cols = c(no_ta, ta_oa),
  values_to = "value",
  names_to = "type") 
```

@@

```{r country_bar_chart_top_20}
top_20_df <- country_scatter_df |>
  filter(cr_year == "2022") |>
  mutate(country_name_fct = forcats::fct_other(country_name, keep = top_20$country_name)) |>
  mutate(country_name_fct = forcats::fct_relevel(country_name_fct, top_20$country_name, "Other")) |>
  group_by(country_name_fct, country_code, cat) |>
  summarise(all = sum(all),
            oa = sum(oa),
            ta_oa = sum(ta_oa), .groups = "drop") |>
  head(20) |>
  mutate(cat_labels = case_when(
   cat == "OECD" ~ glue::glue("<span style = 'color: #3fc1c9'>{country_code}</span>"),
   cat == "BRICS" ~ glue::glue("<span style = 'color: #fc5185'>{country_code}</span>"),
   cat == "Other" ~ glue::glue("<span style = 'color: #e1e1e1'>{country_code}</span>")
  )) |>
  mutate(cat_labels = fct_inorder(cat_labels))

top_20_plot <- top_20_df |>
  mutate(no_ta = oa - ta_oa) |>
  pivot_longer(
  cols = c(no_ta, ta_oa),
  values_to = "value",
  names_to = "type") |>
  ggplot(aes(reorder(as.character(country_code),  value), value / sum(value), fill = type)) +
    geom_bar(stat = "identity", color = "#000000", linewidth = 0.15) +
   scale_fill_manual(
    NULL,
    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
    labels = c(no_ta = "Other", ta_oa = "Agreement"),
    guide = FALSE) +
  coord_flip() +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1L),
    expand = expansion(mult = c(0, 0.05)),
    breaks = scales::breaks_extended(4)
  ) +
  theme_minimal() +
#  theme(axis.text.y = ggtext::element_markdown()) +
  labs(x = NULL, y = "% of Hybrid OA published\nin 2022")

@@

```{r}
country_scatter_22 <- country_scatter_df |>
  filter(cr_year == "2022", !is.na(country_code), all > 5000) |>
  ggplot(aes(all, oa / all)) +
  geom_point(aes(fill = cat), pch = 21, alpha = 0.7) +
  scale_fill_manual(NULL, values = c("OECD" = "#d01c8b", "BRICS" = "#4dac26", "Other" = "#f7f7f7")) +
  geom_smooth(aes(all, oa / all), method = "lm", se = TRUE) +
  scale_size("Agreement", guide = FALSE) +
  ggrepel::geom_text_repel(
    data = filter(country_scatter_df, cr_year == 2022, ta_oa > 2000 | country_code %in% c("CN", "IN", "BR", "RU", "IR")),
    aes(x = all, y = oa / all, label = country_code),
            min.segment.length = .15,
    force = 5,
    segment.color = "grey10",
    segment.size = 0.5,
    segment.alpha = 0.8,
    color = "grey10", 
    size = 3,
    box.padding = 0.3) +
  scale_y_continuous(labels =  scales::percent) +
  scale_x_log10(labels =  scales::number_format(big.mark = ",")) +
  theme_minimal() +
  labs(x = "Total Articles in Hybrid Journals in 2022\n(log-scale)", y = "OA Uptake")

@@

oecd_stats_df <- oecd_stats |>
  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
  group_by(cr_year, cat) |>
  summarise(all = sum(articles, na.rm = TRUE), oa = sum(oa,na.rm = TRUE), ta_oa = sum(ta_oa, na.rm = TRUE)) |>
  mutate(closed = all - oa,
         oa_other = oa - ta_oa) |>
  pivot_longer(cols = c(ta_oa, oa_other, closed), names_to = "oa_type") |>
  mutate(cat = ifelse(is.na(cat), "Other", cat)) |>
  left_join(oecd_by_year, by = "cr_year")

@@

wrap_plots(
  A = country_bar,
  B = country_scatter_22,
  C = top_20_plot,
  design = "
  AC
  BC",
  widths = c(0.7, 0.3),
  heights = c(0.5,0.5)
) &
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10),
    legend.position = "top", 
    legend.justification = "right") &
    plot_annotation(tag_levels = 'A') 
```

@@

```{r country_top_20_plot, fig.asp=0.8, fig.cap="Development of open access in hybrid journals in transformative agreements between 2018 and 2022, by the Top 20 most productive countries in terms of total articles published in the five-years period. Blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established. Country names are represented as ISO two-letter country codes. Facet subheadings show the total number of articles and corresponding market share."}
country_plot
```

>>>>

```{r cr_stats_overview}
cr_stats <- readr::read_csv(here::here("data", "cr_stats.csv"))
cr_stats_plot <- cr_stats |>
  mutate(cat = case_when(
    is.na(src) & oa_status == "gold" ~ "gold",
    src == "hoad" ~ "hoad",
    .default = "other"
  )) |>
  group_by(cr_year, cat) |>
  summarise(n = sum(n_articles)) |>
  mutate(cat = factor(cat, levels = c("hoad", "other", "gold"))) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), n, fill = cat)) +
  geom_bar(
    position = position_stack(reverse = TRUE),
    color = "black",
    stat = "identity"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels =  scales::number_format(big.mark = ",")) +
  scale_fill_manual(
    "Journal",
    values = c(
      "hoad" = "#998ec3",
      "other" = "#f7f7f7",
      "gold" = "#f1a340"
    ),
    labels = c(
      "hoad" = "Hybrid (TA)",
      "other" = "Other",
      "gold" = "Full OA"
    )
  ) +
  theme_minimal() +
  labs(x = "Publication Year", y = "Total Articles") +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title = element_text(size = 10)#,
  #  legend.position = "right",
  #  legend.justification = "right"
  ) 
```

@@

```{r method_fig, fig.asp=0.55, fig.cap="Initial data characteristics. (A) Full open access journals included in transformative agreements by evidence source Directory of Open Access Journals (DOAJ), OpenAlex and the Bielefeld GOLD OA list. (B) Number of articles in Crossref by journal types. The purple bars show the overall article volume of hybrid journals in transformative agreements, which were initially included in the study, in comparision with full open access journals according to OpenAlex. The remainder represents articles in subscription-based journals not covered by transformative agreements."}
(p | cr_stats_plot) + plot_annotation(tag_levels = 'A') +  plot_layout(widths = c(0.6, 0.4))

# Data backup

my_venn_df <- stack(my_venn)
write_csv(my_venn_df, here::here("analysis/fig_data", "fig_2A.csv"))
write_csv(cr_stats_plot$data, here::here("analysis/fig_data", "fig_2B.csv"))
```

@@

oa_jn_box_df <- left_join(jn_hybrid_oa_growth, jn_year_all) |>
  mutate(prop = oa_all / n) |> 
  filter(!cr_year %in%  c("2017", "2023"))
oa_jn_box <- oa_jn_box_df |>
  ungroup() |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  coord_cartesian(ylim = c(0,0.65)) +
  theme_minimal() +
  labs(y = "OA Uptake per Hybrid Journal", x = "Publication Year")
```

@@

layout = "
EEE
ACD
BCD"
wrap_plots(
  E = guide_area(),
  A = oa_ta_plot,
  B = ta_share_plot,
  C = oa_jn_box,
  D = box_ta_prop,
  design = layout,
  widths = c(0.4, 0.3, 0.3),
  heights = c(0.2, 1, .5),
  guides = "collect") &
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
 #   panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A')

# Data export
write_csv(oa_ta_plot$data, here::here("analysis/fig_data", "fig_3A.csv"))
write_csv(ta_share_plot$data, here::here("analysis/fig_data", "fig_3B.csv"))
write_csv(oa_jn_box$data, here::here("analysis/fig_data", "fig_3C.csv"))
write_csv(box_ta_prop$data, here::here("analysis/fig_data", "fig_3D.csv"))
```

@@

publisher_summary_df <- inner_join(oa_publisher, all_publisher) |>
  mutate(prop = value / jn_all) |>
  ungroup()

@@

publisher_ta_prop_plot <- publisher_ta_prop |>
  ungroup() |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_bar(stat = "identity",
           fill = "#56B4E9",
            color = "#000000",
           linewidth = 0.15) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    breaks = c(0, 0.25, 0.5, 0.75),
    limits = c(0, .75),
    labels = scales::percent
  ) +
  facet_wrap( ~ publisher, nrow = 1) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
  theme(
    axis.text.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank()
  )
```

@@

design = "E
A
B
C"
wrap_plots(
  E = guide_area(),
  A = publisher_prop,
  B = publisher_ta_prop_plot,
  C = publisher_boxplot,
  design = design,
  guides = "collect",
  heights = c(0, 0.4, 0.3, 0.7)
) &
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
#    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A')

# Data backup
write_csv(publisher_prop$data, here::here("analysis/fig_data", "fig_4A.csv"))
write_csv(publisher_ta_prop_plot$data, here::here("analysis/fig_data", "fig_4B.csv"))
write_csv(publisher_boxplot$data, here::here("analysis/fig_data", "fig_4C.csv"))
```

@@

```{r subject_plots}
subject_trend_plot <- top_level_by_year_oa |>
  filter(top_level != "Multidisciplinary") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  select(cr_year, top_level, contains("frac"),-oa_frac) |>
  pivot_longer(
    cols = c(ta_oa_frac, no_ta_fractions),
    names_to = "type",
    values_to = "value"
  ) |>
  mutate(prop = value / articles_frac) |>
  ungroup() |>
   ggplot(aes(
    gsub("^20", "'", as.character(cr_year)),
    prop,
    fill = type,
    group = type
  )) +
    geom_area(color = "#000000",
            linewidth = 0.15) +
  facet_wrap( ~ top_level, nrow = 1) +
  scale_fill_manual(
    NULL,
    values = c(
      no_ta_fractions = "#b3b3b3a0",
      ta_oa_frac = "#56B4E9"
    ),
    labels = c(no_ta_fractions = "Other", ta_oa_frac = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    axis.text.x = element_blank()
  )

subject_ta_trend_plot <- top_level_by_year_oa |>
  filter(top_level != "Multidisciplinary") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  mutate(prop = ta_oa_frac / oa_frac) |>
  ungroup() |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_bar(stat = "identity", fill = "#56B4E9", color = "#000000",
            linewidth = 0.15) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
                     labels = scales::percent) +
  facet_wrap(~top_level, nrow = 1) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
  theme(
    axis.text.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank())
subject_box_trend <- subject_df |>
   # with at least one
  select(-ajcs, -subject_area, -esac_publisher) |>
  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
  filter(value > 0) |>
  filter(top_level != "Multidisciplinary", !cr_year %in% c("2017", "2023")) |>
  ungroup() |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE, varwidth = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
  coord_cartesian(ylim = c(0,0.55)) +
  facet_wrap(~top_level, nrow = 1) +
  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
  theme_minimal() +
    theme(strip.background = element_blank(),
    strip.text.x = element_blank()
    )
```

@@

design = "E
A
B
C"
wrap_plots(
  E = guide_area(),
  A = subject_trend_plot,
  B = subject_ta_trend_plot,
  C = subject_box_trend,
  design = design,
  guides = "collect",
  heights = c(0, 0.4, 0.3, 0.7)
) &
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A')

# Data backup

write_csv(subject_trend_plot$data, here::here("analysis/fig_data", "fig_5A.csv"))
write_csv(subject_ta_trend_plot$data, here::here("analysis/fig_data", "fig_5B.csv"))
write_csv(subject_box_trend$data, here::here("analysis/fig_data", "fig_5C.csv"))
```

@@

top_20_plot_df <- 
  country_multiple_df |> 
  mutate(no_ta = oa_total - ta_oa_total) |>
  rename(ta_oa = ta_oa_total) |>
  pivot_longer(
  cols = c(no_ta, ta_oa),
  values_to = "value",
  names_to = "type") |>
  ungroup()
```

@@

```{r country_bar_chart_top_20}
top_20_df <- country_scatter_df |>
  filter(cr_year == "2022") |>
  mutate(country_name_fct = forcats::fct_other(country_name, keep = top_20$country_name)) |>
  mutate(country_name_fct = forcats::fct_relevel(country_name_fct, top_20$country_name, "Other")) |>
  group_by(country_name_fct, country_code, cat) |>
  summarise(all = sum(all),
            oa = sum(oa),
            ta_oa = sum(ta_oa), .groups = "drop") |>
  head(20)

top_20_plot <- top_20_df |>
  mutate(no_ta = oa - ta_oa) |>
  pivot_longer(
  cols = c(no_ta, ta_oa),
  values_to = "value",
  names_to = "type") |>
  ungroup() |>
  ggplot(aes(reorder(as.character(country_code),  value), value / sum(value), fill = type)) +
    geom_bar(stat = "identity", color = "#000000", linewidth = 0.15) +
   scale_fill_manual(
    NULL,
    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
    labels = c(no_ta = "Other", ta_oa = "Agreement"),
    guide = FALSE) +
  coord_flip() +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1L),
    expand = expansion(mult = c(0, 0.05)),
    breaks = scales::breaks_extended(4)
  ) +
  theme_minimal() +
#  theme(axis.text.y = ggtext::element_markdown()) +
  labs(x = NULL, y = "% of Hybrid OA published\nin 2022")

@@

```{r}
country_scatter_22 <- country_scatter_df |>
  filter(cr_year == "2022", !is.na(country_code), all > 5000) |>
  ungroup() |>
  ggplot(aes(all, oa / all)) +
  geom_point(aes(fill = cat), pch = 21, alpha = 0.7) +
  scale_fill_manual(NULL, values = c("OECD" = "#d01c8b", "BRICS" = "#4dac26", "Other" = "#f7f7f7")) +
  geom_smooth(aes(all, oa / all), method = "lm", se = TRUE) +
  scale_size("Agreement", guide = FALSE) +
  ggrepel::geom_text_repel(
    data = filter(country_scatter_df, cr_year == 2022, ta_oa > 2000 | country_code %in% c("CN", "IN", "BR", "RU", "IR")),
    aes(x = all, y = oa / all, label = country_code),
            min.segment.length = .15,
    force = 5,
    segment.color = "grey10",
    segment.size = 0.5,
    segment.alpha = 0.8,
    color = "grey10", 
    size = 3,
    box.padding = 0.3) +
  scale_y_continuous(labels =  scales::percent) +
  scale_x_log10(labels =  scales::number_format(big.mark = ",")) +
  theme_minimal() +
  labs(x = "Total Articles in Hybrid Journals in 2022\n(log-scale)", y = "OA Uptake")

@@

oecd_stats_df <- oecd_stats |>
  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
  group_by(cr_year, cat) |>
  summarise(all = sum(articles, na.rm = TRUE), oa = sum(oa,na.rm = TRUE), ta_oa = sum(ta_oa, na.rm = TRUE)) |>
  mutate(closed = all - oa,
         oa_other = oa - ta_oa) |>
  pivot_longer(cols = c(ta_oa, oa_other, closed), names_to = "oa_type") |>
  mutate(cat = ifelse(is.na(cat), "Other", cat)) |>
  left_join(oecd_by_year, by = "cr_year") |>
  ungroup()

@@

wrap_plots(
  A = country_bar,
  B = country_scatter_22,
  C = top_20_plot,
  design = "
  AC
  BC",
  widths = c(0.7, 0.3),
  heights = c(0.5,0.5)
) &
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10),
    legend.position = "top", 
    legend.justification = "right") &
    plot_annotation(tag_levels = 'A')

# Data backup
write_csv(country_bar$data, here::here("analysis/fig_data", "fig_6A.csv"))
write_csv(country_scatter_22$data, here::here("analysis/fig_data", "fig_6B.csv"))
write_csv(top_20_plot$data, here::here("analysis/fig_data", "fig_6C.csv"))
```

@@

```{r country_top_20_plot, fig.asp=0.8, fig.cap="Development of open access in hybrid journals in transformative agreements between 2018 and 2022, by the Top 20 most productive countries in terms of total articles published in the five-years period. Blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established. Country names are represented as ISO two-letter country codes. Facet subheadings show the total number of articles and corresponding market share."}
country_plot

# Data backup 
write_csv(country_plot$data, here::here("analysis/fig_data", "fig_7.csv"))
```
